\addtocontents{toc}{\protect\vspace{\beforebibskip}} % Place slightly below the rest of the document content in the table


%************************************************
\chapter{Conclusions}
\label{ch:conclusion}
%************************************************

Working in an enterprise environment was very beneficial.
It allowed me to develop a product with much higher quality than the results obtained in previous projects.
It also allowed me to understand how software development is done in the real world, with many intervening parties collaborating for the same goal.

The internship allowed me to improve my skills in two different fields: digital forensics and web development.
The concepts learned from the development of this project allow me to function as a digital forensics professional, 
while the experience gained from the development of the platform largely contributes to my capabilities as a software developer.

My main goal from this internship was to learn skills related to software development.
I obtained a lot of experience working with one of the most used back-end languages, Java, while using Quarkus, a top of the line framework,
and learned a great deal about developing secure platforms.
I also obtained a lot of experience working with React, one of the most popular front-end frameworks, and now I feel capable of producing professional looking user interfaces.

The results obtained from the accomplishment of this internship, the KenSentMe platform, are very gratifying, as I feel that I've exceeded my expectations and produced a high quality product.
These results keep me motivated to further improve my skills and knowledge in software development.

The KenSentMe platform should continue to be included in my work related activities as the code is going to be public on GitHub 
and assuming there's interest in the project, new feature requests and bug reports should pop up and require my assistance.
